MS Prelim Data K23

### 2024-07-18 scripts/first_pass_adapting_K23_analysis_20231031.Rmd
Add files via upload

#findings

@@

@@

@@

#COLLAPSE anxiety dx and antianxiety meds with those plus depression, and anxiety dx OR on anxiety meds into 1 group -->> THIS IS THE MONEY
### Overlap with depression group
```{r overlap_with_dep_group}

@@

@@

```
### Overall Disease by Anxiety Dose *acnp
```{r group_anxietydx_meds_and_dep_comorbidity}
#this collapses anxiety group + meds with those with anxiety group + meds + with/without depression

@@

### Uncinate by Anxiety Dose *acnp?
```{r uncinateByAnxietyDose}

#results: Anxiety dose related to UF_R when covarying for linear effects of age and sex, just barely

@@

########UF R - just looking at uncinate by anxiety dose; anxiety_dose      369.72     187.93   1.967   0.0499 *  *********** (LINEAR AGE)
  anxietyGrouping_UF_R_lm <- lm(UF_R_vol ~ anxiety_dose + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_lm)

  #gam, anxiety_dose    357.0      188.8   1.891   0.0594 .
  anxietyGrouping_UF_R_gam <- gam(UF_R_vol ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_gam)

  #UF _ R anxiety_dose          anxiety_dose      377.35     188.43   2.003    0.046 *********   
  anxietyGrouping_UF_R_by_sex_lm <- lm(UF_R_vol ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_by_sex_lm)

  #gam, anxiety_dose    364.2      189.3   1.924   0.0552 .  
    anxietyGrouping_UF_R_by_sex_gam <- gam(UF_R_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_by_sex_gam)

  #UF L - anxiety_dose     222.193    133.138   1.669    0.096 .  
  anxietyGrouping_UF_L_lm <- lm(UF_L_vol ~ anxiety_dose + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_lm)

   #gam, anxiety_dose    220.1      133.7   1.646    0.101    
  anxietyGrouping_UF_L_gam <- gam(UF_L_vol ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_gam)

  #UF_L anxiety_dose:         anxiety_dose     232.210    133.294   1.742   0.0823 .  
  anxietyGrouping_UF_L_by_sex_lm <- lm(UF_L_vol ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_by_sex_lm)

   #gam, anxiety_dose    229.9      133.9   1.717   0.0868 .  
  anxietyGrouping_UF_L_by_sex_gam <- gam(UF_L_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_by_sex_gam)
```

@@

  #burden of anxiety by network
  #clean up data frame
  df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == c("true_healthy", "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep")) %>% 
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Anxiety_network = proportion_volume_lost_anxiety_net_by_network_size_sum_top_25) %>%
  mutate(Nonanxiety_network = proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75) %>%
  dplyr::select(subject, osex, Diagnosis, Anxiety_network, Nonanxiety_network, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM, -osex, -UF_R_vol, -UF_L_vol, -F_R_vol, -F_L_vol) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#t test
between_dx_t <- t.test(df_empi_dx_network_proportions$proportion_affected[df_empi_dx_network_proportions$Diagnosis == "MS+Anxiety"], df_empi_dx_network_proportions$proportion_affected[df_empi_dx_network_proportions$Diagnosis == "MS-Anxiety"])
summary(between_dx_t)

######means

#mean = 0.5665159
mean_anxiety_dx_anx_net <- mean(df_empi_dx_network_proportions$proportion_affected[which(df_empi_dx_network_proportions$Diagnosis == "MS+Anxiety" & df_empi_dx_network_proportions$Network == "Anxiety_network")])
mean_anxiety_dx_anx_net

#mean = 0.310777
mean_anxiety_dx_nonanx_net <- mean(df_empi_dx_network_proportions$proportion_affected[which(df_empi_dx_network_proportions$Diagnosis == "MS+Anxiety" & df_empi_dx_network_proportions$Network == "Nonanxiety_network")])
mean_anxiety_dx_nonanx_net

#mean =  0.4652665
mean_nonanxiety_dx_anx_net <- mean(df_empi_dx_network_proportions$proportion_affected[which(df_empi_dx_network_proportions$Diagnosis == "MS-Anxiety" & df_empi_dx_network_proportions$Network == "Anxiety_network")])
mean_nonanxiety_dx_anx_net

#mean = 0.2106967
mean_nonanxiety_dx_nonanx_net <- mean(df_empi_dx_network_proportions$proportion_affected[which(df_empi_dx_network_proportions$Diagnosis == "MS-Anxiety" & df_empi_dx_network_proportions$Network == "Nonanxiety_network")])
mean_nonanxiety_dx_nonanx_net

@@

##NetworkNonanxiety_network                      -0.254570   0.013575  92.000000 -18.753  < 2e-16 ***
##DiagnosisMS+Anxiety                             0.108710   0.040130 103.092356   2.709  0.00791 ** 
disease_within_vs_outside_anxiety_net_by_dx_groups_lmer <- lmer(proportion_affected ~ Network*Diagnosis + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_within_vs_outside_anxiety_net_by_dx_groups_lmer)

###do anova instead
##Network             1  3.059  3.0587  82.472  < 2e-16 ***
##Diagnosis           1  0.473  0.4728  12.749 0.000455 ***

disease_within_vs_outside_anxiety_net_by_dx_groups_anova <- aov(proportion_affected ~ Network+Diagnosis + Network:Diagnosis + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_within_vs_outside_anxiety_net_by_dx_groups_anova)

#do lm UF*Diagnosis
 #UF_R_vol -> DiagnosisMS+Anxiety  1418.26     352.04   4.029 8.20e-05 ***
disease_UF_R_vol_lm <- lm(UF_R_vol ~ Diagnosis + osex + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_UF_R_vol_lm)

 #UF_L_vol -> DiagnosisMS+Anxiety   737.97     250.61   2.945  0.00365 ** 
disease_UF_L_vol_lm <- lm(UF_L_vol ~ Diagnosis + osex+PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_UF_L_vol_lm)

#Fornix - Right side but not left
 #F_R_vol -> DiagnosisMS+Anxiety   860.50     368.16   2.337   0.0205 * 
disease_F_R_vol_lm <- lm(F_R_vol ~ Diagnosis + osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_F_R_vol_lm)

 #F_L_vol -> DiagnosisMS+Anxiety  329.724    181.609   1.816   0.0711 . 
disease_F_L_vol_lm <- lm(F_L_vol ~ Diagnosis + osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_F_L_vol_lm)
## net by sex

#simple lmer
#NetworkNonanxiety_network         -0.250361   0.011772  92.000001 -21.268  < 2e-16 ***
  disease_within_vs_outside_anxiety_net_by_sex_diff_lmer <- lmer(proportion_affected ~ Network*osex + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
  summary(disease_within_vs_outside_anxiety_net_by_sex_diff_lmer)

@@

## diagnosis by sex 
 #simple lmer, DiagnosisMS+Anxiety         0.124327   0.040488   3.071  0.00246 **
disease_diagnosis_by_sex_diff_lm <- lm(proportion_affected ~ Diagnosis*osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_diagnosis_by_sex_diff_lm)

@@

#network * diagnosis + sex
 #simple lmer: NetworkNonanxiety_network                      -0.254570   0.013575  92.000000 -18.753  < 2e-16 ***
 ##DiagnosisMS+Anxiety                             0.107571   0.040361 101.820059   2.665  0.00895 ** 
disease_within_vs_outside_anxiety_net_by_dx_groups_sex_diff_lmer <- lmer(proportion_affected ~ Network*Diagnosis + osex + (1|subject)+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_within_vs_outside_anxiety_net_by_dx_groups_sex_diff_lmer)

@@

>>>>

### testing mean UF in depression group  - NS!!!
(previous script (~/BBL/msdepression/scripts/Biological_Psychiatry_Revision_Scripts/msdepression_n380_started_20230927_mm_BP_r1.Rmd)), lines 734 and  779-781
  #bilat: depGroupVar        85.32     206.70   t=0.413    p=0.680    
  disease_mean_UF_vol_lm <- lm(mean_UF_vol ~ depGroupVar + osex+PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
  summary(disease_mean_UF_vol_lm)
#findings

@@

### Define anxiety net by top 25 in both vol/prop
```{r defineanxietynet_with_25_propx25vol}
##### make an array of fascicles that have both substantial overlap with the anxiety network (top 25% vol), and where each fascicle is substantially impacted (top 25% proportion), n = 14
fascicle_names_top_25_vol_x_top_25_prop <- fascicle_names_top_25_vol[fascicle_names_top_25_vol %in% fascicle_names_top_25_prop]

#add vol/prop suffixes
fascicle_names_top_25_vol_x_top_25_prop_PROP_suffix <-fascicle_names_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE)
fascicle_names_top_25_vol_x_top_25_prop_VOL_suffix <-fascicle_names_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

#create fascicle vectors that don't include these special fascicles, n = 63
fascicle_names_NOT_top_25_vol_x_top_25_prop <- fascicle_names_no_cranial_nerves[!(fascicle_names_no_cranial_nerves %in% fascicle_names_top_25_vol_x_top_25_prop)]

#add vol/prop suffixes
fascicle_names_NOT_top_25_vol_x_top_25_prop_PROP_suffix <- fascicle_names_NOT_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE)
fascicle_names_NOT_top_25_vol_x_top_25_prop_VOL_suffix <- fascicle_names_NOT_top_25_vol_x_top_25_prop%>% gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

#anxiety25_propxvol overlap total, sum of these two is 1545449 (and this is total_fascicle_volume)
anxiety_top_25_vol_x_top_25_prop_vol <- sum(dep_net_overlap_from_bp_paper$num_voxels_total[dep_net_overlap_from_bp_paper$fascicle %in% fascicle_names_top_25_vol_x_top_25_prop]) #vol = 181869

#nonanxiety25_propxvol overlap total
anxiety_NOT_top_25_vol_x_top_25_prop_vol <- sum(dep_net_overlap_from_bp_paper$num_voxels_total[dep_net_overlap_from_bp_paper$fascicle %in% fascicle_names_NOT_top_25_vol_x_top_25_prop]) #vol = 1363580

#check sums
if ((anxiety_top_25_vol_x_top_25_prop_vol + anxiety_NOT_top_25_vol_x_top_25_prop_vol) == total_fascicle_volume) {
  print("anxiety + nonanxiety net = total fascicle volume")
} else{
  print("something is wrong with counts")
}

@@

    ##sum of total volume lost in anxiety 25 volxanxiety25prop groups, and proportion of volume lost in network (volume lost/network volume size)
  mutate(sum_fascicle_names_top_25_vol_x_top_25_prop = 
           rowSums(across(all_of(fascicle_names_top_25_vol_x_top_25_prop_VOL_suffix)))) %>%
  mutate(proportion_fascicle_vol_lost_top_25_vol_x_top_25_prop = sum_fascicle_names_top_25_vol_x_top_25_prop/anxiety_top_25_vol_x_top_25_prop_vol) %>%

    ##sum of total volume lost in remaining 63 fascicles NOT in anxiety 25 volxanxiety25prop groups, and proportion of volume lost in network (volume lost/network volume size)
  mutate(sum_fascicle_names_NOT_top_25_vol_x_top_25_prop = 
           rowSums(across(all_of(fascicle_names_NOT_top_25_vol_x_top_25_prop_VOL_suffix)))) %>%
  mutate(proportion_fascicle_vol_lost_NOT_top_25_vol_x_top_25_prop = sum_fascicle_names_NOT_top_25_vol_x_top_25_prop/anxiety_NOT_top_25_vol_x_top_25_prop_vol) %>%

   ##mean_uncinate)
  mutate(mean_UF_vol = (UF_L_vol + UF_R_vol)/2) %>%
  mutate(mean_UF_prop = (UF_L_prop + UF_R_prop)/2) %>%

@@

#COLLAPSE anxiety dx and antianxiety meds with those plus depression, and anxiety dx OR on anxiety meds into 1 group -->> THIS IS THE MONEY
### Overlap with depression group
```{r overlap_with_dep_group}
#### 53.4% overlap total

@@

#"Percent overlap 53.4%"
print(paste0("Percent overlap ", round(sum(df_disease_burden_unc_and_anx_net_collapsed_variables$EMPI_in_dep_group)/length(df_disease_burden_unc_and_anx_net_collapsed_variables$EMPI), 3) * 100, "%"))

@@

### Overall Disease by Anxiety Dose *acnp
```{r group_anxietydx_meds_and_dep_comorbidity}
#this collapses anxiety group + meds with those with anxiety group + meds + with/without depression

@@

### Lesion volumes 
#### Diagnosis 
  - sick > healthy in vol of lesions
```{r lesion_volumes_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, volume_of_mimosa_lesions) %>%
  ungroup()

##### healthy vs sick

#lm lesion volumes; DiagnosisMS+Anxiety  6310.99    2143.87   2.944 0.003694 ** 
volume_of_mimosa_lesions_lm <- lm(volume_of_mimosa_lesions ~ Diagnosis + osex + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(volume_of_mimosa_lesions_lm)

#gam lesion volumes; DiagnosisMS+Anxiety   6973.5     2014.7   3.461 0.000681 ***
volume_of_mimosa_lesions_gam <- gam(volume_of_mimosa_lesions ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(volume_of_mimosa_lesions_gam)
```

#### Anxiety dose
  - anxiety dose correlates with disease burden (volume of lesions)
```{r lesion_volumes_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != 3)

#dose; anxiety_dose     3028.07    1014.45   2.985 0.003026 **

  anxiety_dose_volume_of_mimosa_lesions_by_sex_lm <- lm(volume_of_mimosa_lesions ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_lm)

 #bilat anxiety dose + sex + age: anxiety_dose   3304.7      985.8   3.352 0.000886 ***

  anxiety_dose_volume_of_mimosa_lesions_by_sex_gam <- gam(volume_of_mimosa_lesions ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

### Uncinate **ACNP**

#### Demographics
```{r demo_uncinate}

###### all groups, no unclassified
make_demographics_table_ms_anxiety_simple(df_demo_and_fascicles %>% filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified"))

###### just healthy and sick
make_demographics_table_ms_anxiety_simple(df_demo_and_fascicles %>% filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds"))

```
#### Uncinate by diagnosis 
```{r uncinate_by_dx}

### Results: LM Mean UF_R & L, age, sex cov, DiagnosisMS+Anxiety   621.95     276.11   2.252   0.0256 * 
#define group
  df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol, mean_UF_vol) %>%
  ungroup()

#do lm UF*Diagnosis
 #UF_R_vol -> DiagnosisMS+Anxiety   793.14     361.94   2.191   0.0298 *   
disease_UF_R_vol_lm <- lm(UF_R_vol ~ Diagnosis + osex + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_UF_R_vol_lm)

 #UF_L_vol ->DiagnosisMS+Anxiety   450.75     261.20   1.726   0.0862 .    
disease_UF_L_vol_lm <- lm(UF_L_vol ~ Diagnosis + osex+PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_UF_L_vol_lm)

#bilat: DiagnosisMS+Anxiety   621.95     276.11   2.252   0.0256 *  
disease_mean_UF_vol_lm <- lm(mean_UF_vol ~ Diagnosis + osex+PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_mean_UF_vol_lm)

#gam
#UF_R_vol -> DiagnosisMS+Anxiety   795.60     366.58   2.170   0.0314 *   
disease_UF_R_vol_gam <- gam(UF_R_vol ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(disease_UF_R_vol_gam)

 #UF_L_vol -> DiagnosisMS+Anxiety   420.24     262.19   1.603    0.111   
disease_UF_L_vol_gam <- gam(UF_L_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(disease_UF_L_vol_gam)

#bilateral UF: DiagnosisMS+Anxiety   607.92     279.34   2.176   0.0309 *  
disease_mean_UF_vol_gam <- gam(mean_UF_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(disease_mean_UF_vol_gam)

#Fornix - Right side but not left
 #F_R_vol -> DiagnosisMS+Anxiety   860.50     368.16   2.337   0.0205 * 
#disease_F_R_vol_lm <- lm(F_R_vol ~ Diagnosis + osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
#summary(disease_F_R_vol_lm)

 #F_L_vol -> DiagnosisMS+Anxiety  329.724    181.609   1.816   0.0711 . 
#disease_F_L_vol_lm <- lm(F_L_vol ~ Diagnosis + osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
#summary(disease_F_L_vol_lm)
## net by sex 
```

#### Uncinate by Anxiety Dose 
```{r uncinateByAnxietyDose}

#results: Anxiety dose related to UF_R when covarying for linear effects of age and sex, just barely
# Additionally - taking the mean of both UF and UR, LM for bilat anxiety dose + sex + age: anxiety_dose most sig P= 0.0367 *

@@

#lm

  #R anxiety dose + age ; anxiety_dose      369.72     187.93   1.967   0.0499 *  ***********
  anxietyGrouping_UF_R_lm <- lm(UF_R_vol ~ anxiety_dose + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_lm)

   #R *anxiety dose + osex + age anxiety_dose          anxiety_dose      377.35     188.43   2.003    0.046 *********   
  anxietyGrouping_UF_R_by_sex_lm <- lm(UF_R_vol ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_by_sex_lm)

   #UF L - anxiety_dose + age    222.193    133.138   1.669    0.096 .  
  anxietyGrouping_UF_L_lm <- lm(UF_L_vol ~ anxiety_dose + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_lm)

   #UF_L anxiety_dose + sex + age:         anxiety_dose     232.210    133.294   1.742   0.0823 .  
  anxietyGrouping_UF_L_by_sex_lm <- lm(UF_L_vol ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_by_sex_lm)

  #bilat anxiety dose + age: anxiety_dose     295.957    145.116   2.039   0.0421 *  
  anxietyGrouping_mean_UF_lm <- lm(mean_UF_vol ~ anxiety_dose + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_lm)

  #bilat anxiety dose + sex + age: anxiety_dose      304.78     145.39   2.096   0.0367 *  
  anxietyGrouping_mean_UF_by_sex_lm <- lm(mean_UF_vol ~ anxiety_dose + osex + PAT_AGE_AT_EXAM, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_lm)

#gam

  #UF _ R  anxiety_dose + age   357.0      188.8   1.891   0.0594 .
  anxietyGrouping_UF_R_gam <- gam(UF_R_vol ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_gam)

 #R anxiety dose + sex + age: anxiety_dose    364.2      189.3   1.924   0.0552 .

    anxietyGrouping_UF_R_by_sex_gam <- gam(UF_R_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_R_by_sex_gam)

   #UF_L, anxiety_dose + age    220.1      133.7   1.646    0.101    
  anxietyGrouping_UF_L_gam <- gam(UF_L_vol ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_gam)

   #UF_L anxiety_dose + sex + age    229.9      133.9   1.717   0.0868 .  
  anxietyGrouping_UF_L_by_sex_gam <- gam(UF_L_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_UF_L_by_sex_gam)

  #bilat

  #bilat  anxiety_dose + age: anxiety_dose    288.5      145.7    1.98   0.0484 *  
  anxietyGrouping_mean_UF_gam <- gam(mean_UF_vol ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_gam)

 #bilat anxiety dose + sex + age: anxiety_dose    297.0      146.0   2.034   0.0426 *

    anxietyGrouping_mean_UF_by_sex_gam <- gam(mean_UF_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam)
```

@@

  #burden of anxiety by network
  #clean up data frame
  df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Anxiety_network = proportion_volume_lost_anxiety_net_by_network_size_sum_top_25) %>%
  mutate(Nonanxiety_network = proportion_volume_lost_nonanxiety_net_by_network_size_sum_bottom_75) %>%
  dplyr::select(subject, osex, Diagnosis, Anxiety_network, Nonanxiety_network, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis, -PAT_AGE_AT_EXAM, -osex, -UF_R_vol, -UF_L_vol, -F_R_vol, -F_L_vol) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

@@

#NetworkNonanxiety_network                      -0.246157   0.009900 173.000001 -24.864  < 2e-16 ***
#DiagnosisMS+Anxiety                             0.075874   0.030315 194.726132   2.503 0.013142 *

disease_within_vs_outside_anxiety_net_by_dx_groups_lmer <- lmer(proportion_affected ~ Network*Diagnosis + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_within_vs_outside_anxiety_net_by_dx_groups_lmer)

#Network confit:
#NetworkNonanxiety_network                     -0.265556308 -0.226758160
#DiagnosisMS+Anxiety                            0.016671783  0.135076698
confint(disease_within_vs_outside_anxiety_net_by_dx_groups_lmer)

#simple lmer
#NetworkNonanxiety_network         -0.248004   0.008994 173.000001 -27.573  < 2e-16 ***
  disease_within_vs_outside_anxiety_net_by_sex_diff_lmer <- lmer(proportion_affected ~ Network*osex + (1|subject) + PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
  summary(disease_within_vs_outside_anxiety_net_by_sex_diff_lmer)

@@

## diagnosis by sex 
 #simple lmer, DiagnosisMS+Anxiety         0.102015   0.030998   3.291   0.0011 ** 
disease_diagnosis_by_sex_diff_lm <- lm(proportion_affected ~ Diagnosis*osex+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_diagnosis_by_sex_diff_lm)

@@

#network * diagnosis + sex
#NetworkNonanxiety_network                      -0.246157   0.009900 173.000001 -24.864  < 2e-16 ***
#DiagnosisMS+Anxiety                             0.076902   0.030464 193.366257   2.524 0.012394 * 
disease_within_vs_outside_anxiety_net_by_dx_groups_sex_diff_lmer <- lmer(proportion_affected ~ Network*Diagnosis + osex + (1|subject)+ PAT_AGE_AT_EXAM, data = df_empi_dx_network_proportions)
summary(disease_within_vs_outside_anxiety_net_by_dx_groups_sex_diff_lmer)

@@

## Anxiety net defined as (25 vol x 25 prop anxiety net) -> no better, don't use
### Anxiety and Non anxiety net by dose 
```{r anxiety_and_nonanxiety_net_by_dose}

### Will be using gam + s(age) and osex because this was giving me most robust findings previously.
### I know that we already have an effect in across whole brain. Now i want to check if it differs within new anxiety net vs outside

df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>% 
  dplyr::select(proportion_volume_lost_per_total_network_size_sum, proportion_fascicle_volume_lost_top_25_vol_x_top_25_prop, proportion_fascicle_vol_lost_NOT_top_25_vol_x_top_25_prop, UF_L_prop, UF_R_prop, oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed, osex, anxiety_dose, PAT_AGE_AT_EXAM)

## gam, check original finding still there, anxiety_dose  0.03692    0.01496   2.468    0.014 *  
   anxietyGrouping_cov_sex_gam <- gam(proportion_volume_lost_per_total_network_size_sum ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_disease_burden_unc_and_anx_net_collapsed_variables)
  summary(anxietyGrouping_cov_sex_gam)

  #gam, anxiety_dose, anxiety net, anxiety_dose  0.027830   0.015810   1.760   0.0792 . 
   anxietyGrouping_cov_sex_anxiety_net_gam <- gam(proportion_fascicle_volume_lost_top_25_vol_x_top_25_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_disease_burden_unc_and_anx_net_collapsed_variables)
  summary(anxietyGrouping_cov_sex_anxiety_net_gam)

    #gam, anxiety_dose, non anxiety net; anxiety_dose  0.03814    0.01500   2.543   0.0114 *  
   anxietyGrouping_cov_sex_nonanxiety_net_gam <- gam(proportion_fascicle_vol_lost_NOT_top_25_vol_x_top_25_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_disease_burden_unc_and_anx_net_collapsed_variables)
  summary(anxietyGrouping_cov_sex_nonanxiety_net_gam)

## take home: comparing healthy to has anxiety AND meds, more disease within vs outside net, anxiety vs non, no interaction.
## take home #2 : More disease in R & L uncinate in anxious versus nonanxious
## take home #3 : About the same results as using the previously described anxiety net (rather than the vol 25 x 25 prop)
